# TUTORIAL
## Mapping Voxel-based niftis to ROI-based surface data
### prepare environment and load libraries
Make sure you have installed the latest version of the connectome workbench, which you can download here:
https://www.humanconnectome.org/software/connectome-workbench

```{r message=FALSE, warning=FALSE}
#### load required libaries ####
require(neurobase)
require(ciftiTools)
require(plyr)
require(dplyr)

#### set path to workbench (example for a MacOS environment) ####
ciftiTools.setOption('wb_path', '/Applications/workbench/bin_macosx64')

```

### extract ROI-based data
For this example, we will extract ROI based data from a tau-PET image
using the 17 network version of the Schaefer 200 ROI atlas.
The atlas including the reference (Schaefer et al., Cerebral Cortex, 2018) can be found here:
https://github.com/ThomasYeoLab/CBIG/tree/master/stable_projects/brain_parcellation/Schaefer2018_LocalGlobal

### set tutorial path
ste the path to your downloaded tutorial
```{r}
sdir = "/Users/isd_neuroimaging/Library/Mobile Documents/com~apple~CloudDocs/Github/Tutorials/Connectome_workbench/Map_nifti_to_ROIs/"
```

### read in images
#### read in atlas 
```{r}

# load the atlas
atlas_readin <- neurobase::readnii(
  file.path(sdir, "/Example_data/rSchaefer2018_200Parcels_17Networks_order.nii.gz")
)
neurobase::ortho2(atlas_readin)

```


#### read in the PET image 
```{r}

# load the atlas
pet_readin <- neurobase::readnii(
  file.path(sdir, "/Example_data/Example_tau.nii.gz")
)

neurobase::ortho2(pet_readin)

```

#### check overlap between atlas and pet image
```{r}
# check overlay
neurobase::ortho2(pet_readin, atlas_readin)

# check if dimensions match
dim(pet_readin) == dim(atlas_readin)
```


### extract ROI-based mean values

```{r}
pet_atlas_based = 
  data.frame(pet = as.vector(pet_readin@.Data),
             atlas_idx = as.vector(atlas_readin@.Data)) %>% 
  filter(atlas_idx !=0) %>% 
  group_by(atlas_idx) %>% 
  summarize(mean = mean(pet))
head(pet_atlas_based)

# show distribution of SUVR data
hist(pet_atlas_based$mean)

```

### map to xifti file
```{r}

# read in atlas
atlas_dscalar <- read_xifti(file.path(sdir, "/Example_data/Schaefer2018_200Parcels_17Networks_order.dscalar.nii"))
atlas_dscalar_vec <- c(as.matrix(atlas_dscalar))

# determine values to map to surface
vector_to_map = pet_atlas_based$mean

# map ROI-based PET data to cifti
xii_to_map <- c(NA, vector_to_map)[atlas_dscalar_vec + 1]
xii1 <- select_xifti(atlas_dscalar, 1)
xii_to_map <- newdata_xifti(xii1, xii_to_map)

# view cifti directly in R
view_cifti_surface(xii_to_map, colors = "magma", zlim = range(vector_to_map),
                   borders = T)
  
```
### write cifti to disk
```{r}
ciftiTools::write_cifti(xii_to_map, file.path(sdir, "/Example_surface"))

```



